Report on Late Winter2013 Production: Image Differencing

### 2013-03-22 W13report.tex
Laid out the long list of conclusions.

We also list the ratio of negatively detected to
positively detected DiaSources.  That this ratio is significantly
greater than 1.0 in most cases suggests that there is an
oversubtraction of the background in the process, which biases this
ratio (see Section~\ref{sec-bg}).  The optimal configuartions yielding
these false detection rates are given in Table~\ref{tab-bestconfig10}.

We have examined the distribution of these false detectiosn across the
image, and find no significant overdensities near the borders of the
images, which would be expected in cases of under or over--fitting of
the various spatial models (astrometry, backgrounds, Psf--matching
kernels).  Figure~\ref{edgedist} provides the distribution of
distances from the edge of the sensor for all false detections in {\it
  blue}, with a random sample of points in the image given in {\it
  green}.  Uncertainties are plotted using the square root of the
number of points in each bin, and the histograms are normalized to
provide a probability density.  The takeaway is that we find no
significant correlation of false detetions with the boundaries of the
fitting functions.

\section{False Detecion Dependencies}

After finding the configurations that yielded the minimal numbers of
false detections, we then perturbed the solutions to examine how the
numbers of false detections scaled with different effects.  This
included, in order of importance for this production: the difference
image detection threshold; the number of KernelCandidates going into
the spatial model; and astrometric registration errors.  We also
undertake a theoretical analysis of the number of false detections
expected in a 4k x 4k image as a function fo detection threshold, and
examine the requirements this sets on our ability to model the
variance and backgrounds in the images.  We compare the emperical
variation of false detections with detection threshold, and show that
our data are consistent with this.

@@

The variance in the prefilterd and postfiltered images is likely to be
different due to the different types of operations that have happened
on the images.  We quantify this by taking the postfiltered images,
immediately before the detection step (i.e. $D'$), and compare the
variance plane to that coming from the prefiltered image.  For
v6866601, which is a deconvolution in the postfiltered production, we
find that the variance is 70\% higher when postfiltering.  This is
consistent with the expectation that deconvolution increases the noise
properties of the image.

For v6866602 and v6866603, we find that the variance is 3--4\% and
2--3\% higher in the {\it prefiltered} data, respectively.  The
primary culprit for this is that the prefiltered data will have gone
through 1 less convolution, meaning less of the variance will have
been lost to covariance.  In this case, the variance plane will be
closer to the truth than in the postfiltered data, but still not
exact.  Assuming that most of the covariance happens during
Psf--matching and Psf--filtering (the astrometric warping kernel is
designed to preserve the nosie properties of the image, although this
is not an exact conservation in practice), the variance plane of the
prefiltered data should itself be $\sim$3\% lower than the emperical
variance in the image plane.  {\it TEST ME}.

@@

Pre--filtering is clearly preferred when the traditional process would
lead to deconvolution, or sharpening, of the data.  When the standard
processing results in a convolution, or smoothing of the data,
pre--filtering provides fewer false detections.  As a trade--off, the
measurement of these false detetions is not yet well defined.

@@

\begin{table}
\centering
\begin{tabular}{clcc}
\hline
\multicolumn{4}{|c|}{Variance Ratios} \\
\hline
Visit    & Prefilter & Mean Ratio & RMS Ratio \\
\hline
v6866601 & True      & 1.009      & 0.002     \\
         & False     & 1.005      & 0.001     \\
v6866602 & True      & 1.003      & 0.0004    \\
         & False     & 1.005      & 0.0007    \\
v6866603 & True      & 1.004      & 0.0003    \\
         & False     & 1.008      & 0.002     \\
\end{tabular}
\caption{Ratio of the emperical RMS in the difference images,
  calculated via the interquartile range, to the square root of the
  median of the variance plane.  In all cases the variance plane
  represents an {\it underestimate} of the true variance in the
  images.  We report the mean and RMS across all sensors.
}
\end{table}

>>>>

We also list the ratio of negatively detected to
positively detected DiaSources.  That this ratio is significantly
greater than 1.0 in most cases suggests that there is an
over-subtraction of the background in the process, which biases this
ratio (see Section~\ref{sec-bg}).  The optimal configurations yielding
these false detection rates are given in Table~\ref{tab-bestconfig10}.

We have examined the distribution of these false detections across the
image, and find no significant overdensities near the borders of the
images, which would be expected in cases of under or over--fitting of
the various spatial models (astrometry, backgrounds, Psf--matching
kernels).  Figure~\ref{edgedist} provides the distribution of
distances from the edge of the sensor for all false detections in {\it
  blue}, with a random sample of points in the image given in {\it
  green}.  Uncertainties are plotted using the square root of the
number of points in each bin, and the histograms are normalized to
provide a probability density.  The takeaway is that we find no
significant correlation of false detections with the boundaries of the
fitting functions.

\section{False Detection Dependencies}

After finding the configurations that yielded the minimal numbers of
false detections, we then perturbed the solutions to examine how the
numbers of false detections scaled with different effects.  This
included, in order of importance for this production: the difference
image detection threshold; the number of KernelCandidates going into
the spatial model; and astrometric registration errors.  We also
undertake a theoretical analysis of the number of false detections
expected in a 4k x 4k image as a function of detection threshold, and
examine the requirements this sets on our ability to model the
variance and backgrounds in the images.  We compare the empirical
variation of false detections with detection threshold, and show that
our data are consistent with this.

@@

The variance in the prefiltered and postfiltered images is likely to be
different due to the different types of operations that have happened
on the images.  We quantify this by taking the postfiltered images,
immediately before the detection step (i.e. $D'$), and compare the
variance plane to that coming from the prefiltered image.  For
v6866601, which is a deconvolution in the postfiltered production, we
find that the variance is 70\% higher when postfiltering.  This is
consistent with the expectation that deconvolution increases the noise
properties of the image.

For v6866602 and v6866603, we find that the variance is 3--4\% and
2--3\% higher in the {\it prefiltered} data, respectively.  The
primary culprit for this is that the prefiltered data will have gone
through 1 less convolution, meaning less of the variance will have
been lost to covariance.  In this case, the variance plane will be
closer to the truth than in the postfiltered data, but still not
exact.  Assuming that most of the covariance happens during
Psf--matching and Psf--filtering (the astrometric warping kernel is
designed to preserve the noise properties of the image, although this
is not an exact conservation in practice), the variance plane of the
prefiltered data should itself be $\sim$3\% lower than the empirical
variance in the image plane.  {\it TEST ME}.

@@

Pre--filtering is clearly preferred when the traditional process would
lead to deconvolution, or sharpening, of the data.  When the standard
processing results in a convolution, or smoothing of the data,
pre--filtering provides fewer false detections.  As a trade--off, the
measurement of these false detections is not yet well defined.

\begin{itemize}

\item We find that the overall number of false detections is not
  strongly sensitive to bulk background misestimation
  (Figure~\ref{XXX});

\item However, the ratio of positive to negative false detections is
  strongly dependent on background fitting, with {\bf XXX}
  (Figure~\ref{XXX}).

\item At the 5--sigma detection threshold, for the range of seeings
  considered in this production, the theoretical numbers of false
  detections per 4k x 4k sensor numbers 10+.  This function is steep;
  by increasing the detection threshold to 5.X sigma this number may
  be lowered to less than 1 per sensor.

\item Due to the steepness of this function, the number of false
  detections is strongly dependent on the variance being correct, at
  the 5--sigma detection threshold.  When detecting at 5--sigma,
  having the variance underestimated by 1\% leads to an increase
  in false positives of {\bf XXX} (Figure~\ref{}).

\item An empirical computation of the variance from the image plane,
  compared to the median propagated variance in the variance plane,
  indicates that we consistently underestimate the variance by $1-2\%$
  (Table~\ref{tab-variance1}).  A proper tracking of covariance during
  the multiple image convolutions that occur may alleviate this
  problem.

\item When convolving to Psf--match, variance planes of postfiltered
  detection images have {\it lower} variance than their prefiltered
  counterparts by $2-3\%$ (Table~\ref{tab-variance2}).  These images
  have gone through one additional convolution compared to the
  prefiltered data.

\item When deconvolving, the postfiltered variance is 70\% {\it
  higher} at the time of detection compared to the prefiltered data
  (Table~\ref{tab-variance2}).  This is consistent with the
  understanding that deconvolution increases high frequency noise in
  the images, and provides a {\it lower} detection efficiency in
  deconvolved data.  Prefiltering is clearly preferred in the case of
  deconvolution.

\item The false detection rate is seen to scale with detection
  threshold in a manner consistent with theory (Figure~\ref{}).

\item Several statistics that we are able to calculate at the time of
  image subtraction are shown to correlate strongly with the numbers
  of false detections, providing real--time quality estimates
  (Figures~\ref{}).

\item The numbers of false detections that associate with Sources is
  shown to correlate with the amount of smoothing that occurs in the
  Psf matching, in the sense that more smoothing leads to fewer false
  detections (Figure~\ref{}).  In this sense, there is a strong
  preference for running in prefiltered mode.

\item Measurements on prefiltered difference images contain less
  information than in postfiltered difference images, because the
  paradigm of detection on filtered data but measurement on the
  unfiltered data is broken.  Thus while we detect on average fewer
  false positives in prefiltered data, we are (so far) able to
  characterize them less well.  This trade--off is clearly something
  to explore in future work.

\end{itemize}

@@

\begin{table}
\centering
\begin{tabular}{clcc}
\hline
\multicolumn{4}{|c|}{Variance Ratios} \\
\hline
Visit    & Prefilter & Mean Ratio & RMS Ratio \\
\hline
v6866601 & True      & 1.019      & 0.005    \\
         & False     & 1.011      & 0.002    \\
v6866602 & True      & 1.007      & 0.001    \\
         & False     & 1.010      & 0.002    \\
v6866603 & True      & 1.008      & 0.001    \\
         & False     & 1.015      & 0.003    \\
\end{tabular}
\caption{Ratio of the empirical variance in the difference images,
  calculated via (0.741 times the interquartile range)$^2$, to the
  median of the variance plane.  In all cases the variance plane
  represents an {\it underestimate} of the true variance in the
  images.  We report the mean and RMS across all sensors.  }
\label{tab-variance1}
\end{table}

\begin{table}
\centering
\begin{tabular}{ccc}
\hline
\multicolumn{3}{|c|}{Variance Ratios} \\
\hline
Visit    & Mean Postfilter/Prefilter & RMS Postfilter/Prefilter \\
\hline
v6866601 & 1.72      & 0.04     \\
v6866602 & 0.967     & 0.001    \\
v6866603 & 0.975     & 0.002    \\
\end{tabular}
\caption{Ratios of the variance planes immediately before detection
  $D'$.  Ratios are in the sense of the variance of the postfiltered
  image divided by the variance of the prefiltered image.  We report
  the mean and RMS across all sensors.}
\label{tab-variance2}
\end{table}
